### 2022-03-18 pre/pre.tex
more

@@

\looseness=-1
An {\em equation} $s \eq t$ corresponds to an unordered pair of terms. A {\em
literal} $l$ is an equation $s \eq t$ or its negation (disequation) $s \noteq
t$. A clause $C$ is a finite multiset of literals, interpreted and written
disjunctively:\ $l_1 \llor \cdots \llor l_n$. Free variables of clauses are
implicitly universally quantified. Note that $\llor$ denoting disjunction in
clauses is different from the formula-building $\ior$. In general,
formula-building logical connectives are written in boldface.
In standard, non-clausal first-order logic, an atom is either
predicate or equational. For uniformity, and to
stay close to the implementation, we encode predicate atoms as equations with
$\itrue$. Negative literals are encoded as disequations. For example,
$\cst{even}(x)$ is encoded as $\poslit{\cst{even}(x)}$, and
$\neg\,\cst{even}(x)$ is encoded as $\neglit{\cst{even}(x)}$. To lighten the notation,
we sometimes write the predicate literals in non-encoded form.

Applying substitution to a literal is reduced to applying it to both sides of
the (dis)equa\-tion and it is extended pointwise to clauses. It is sometimes
also written in postfix notation as $l\sigma$ and $C\sigma$. We say
$\sigmacl{C}$ is a ground instance of $C$ if $\sigma$ is a grounding
substitution.

>>>>

Some of the approaches described in this thesis are based on {\em rank-1
polymorphism} \cite{bp-13-tff1,ksr-16-th1}. This is an extension of simply typed
logic which adds type variables to the definition of types and type arguments to
polymorphic constants (Chapter \ref{ch:bools}). It further
requires that type arguments are instantiated with concrete, non-quantified
types. In clausal structure it allows (implicit)
quantification over types only on top-level and not inside literals.

@@

\looseness=-1
An {\em equation} $s \eq t$ corresponds to an unordered pair of terms. A {\em
literal} $l$ is an equation $s \eq t$ or its negation (disequation) $s \noteq
t$. A clause $C$ is a finite multiset of literals, interpreted and written
disjunctively:\ $l_1 \llor \cdots \llor l_n$. Free variables of clauses are
implicitly universally quantified. Note that $\llor$ denoting disjunction in
clauses is different from the formula-building $\ior$. In general,
formula-building logical connectives are written in boldface. In standard,
non-clausal first-order logic, an atom is either predicate or equational. For
uniformity, and to stay close to the implementation, we encode predicate atoms
as equations with $\itrue$. Negative literals are encoded as disequations. For
example, $\cst{even}(x)$ is encoded as $\poslit{\cst{even}(x)}$, and
$\neg\,\cst{even}(x)$ is encoded as $\neglit{\cst{even}(x)}$. To lighten the
notation, we sometimes write the predicate literals in non-encoded form.
Applying a substitution to a literal is reduced to applying it to both sides of
the (dis)equa\-tion and it is extended pointwise to clauses. We say
$\sigmacl{C}$ is a ground instance of $C$ if $\sigma$ is a grounding
substitution.
\pagebreak[2]

### 2022-03-18 satfol/satfol.tex
more

We assume the same notion of substitution introduced in Sect.~\ref{sec:pre:fol}.
We extend the notation by writing a substitution $\{x_1 \mapsto t_1,
\ldots,\allowbreak x_n \mapsto t_n\}$ shortly as $\{ \tupleempty{x} \mapsto \tupleempty{t} \}$.
A substitution is a \emph{variable renaming} if it is a
bijection from a set of variables to a set of variables.

@@

We assume the natural extensions of domain, valuation, interpretation and model
(as defined by Fitting \cite{mf-1996-fol}) from unsorted to many-sorted logic.
We also assume the notion of normal models from Sect.~\ref{sec:pre:fol} (with
corresponding notation) and extend it with the notion of \emph{cannonical
models}: A canonical model $\ourmodel$ is a normal model such that for every
element $d$ in one of $\ourmodel$'s domains there exists a ground term $t$ such
that $\ourmodel$ interprets $t$ as $d$. Canonical models generalize Herbrand
models \cite{mf-1996-fol} to first-order logic with equality: If a clause set is satisfiable, then
it is satisfiable in a canonical model.

>>>>

We assume the same notion of substitution introduced in Sect.~\ref{sec:pre:fol}.
We extend the notation by writing a substitution $\{x_1 \mapsto t_1,
\ldots,\allowbreak x_n \mapsto t_n\}$ shortly as $\{ \tupleempty{x} \mapsto \tupleempty{t} \}$.
A substitution is a \emph{variable renaming} if it is a
bijection from a set of variables to a set of variables. Iterated, $i$-fold application 
of nonidempotent application $\sigma$ is denoted by $\sigma^i$.

@@

We assume the natural extensions of domain, valuation, interpretation and model
(as defined by Fitting \cite{mf-1996-fol}) from unsorted to many-sorted logic.
We also assume the notion of normal models from Sect.~\ref{sec:pre:fol} (with
corresponding notation) and extend it with the notion of \emph{cannonical
models}: A canonical model $\ourmodel$ is a normal model such that for every
element $d$ in one of $\ourmodel$'s domains there exists a ground term $t$ such
that $\ourmodel$ interprets $t$ as $d$. 
We also make use of Herbrand models and Herbrand's theorem \cite[Sect.~5.4]{mf-1996-fol}.
Canonical models generalize Herbrand
models to first-order logic with equality: If a clause set is satisfiable, then
it is satisfiable in a canonical model.
